### 2020-06-25 TCS_version/07-discussion.tex
altri piccoli cambiamenti
embedding encoding translation

% !TEX root = ./main_RS2L_LNCS.tex
%\section{Discussion}

@@

Our encoding increases the expressivity of  RS  concerning:
%the behaviour of the context,
the possibility of alternative behaviour of
mutated entities, and  the communication between two different Reaction Systems.
It is important to note that, when the context is deterministic, our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one state is reachable, 
as the dynamics is totally deterministic.

@@

\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline
$rs1$  & $rs2$ \\
\hline
%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
$a_1=(s,\ , x)$ & $a_2=(y,\  ,s)$\\
%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
\hline 
\end{tabular}
\caption{The two Reaction Systems $rs1$ and $rs2$.}
\label{tab:rules}
\end{table}
 \noindent
Now, we set our example such that the two contexts, for  $rs1$ and $rs2$, do not provide any entities.
We also assume that entity $s$ in $rs1$ is provided by $rs2$, as $rs2$ produces a quantity of $s$ that is enough for $rs1$ and $rs2$.
For technical reasons, we can not use the same name for  $s$ in both the two RSs, then we use the name $ss$ in $rs2$.
%
We need to modify our translation technique to suite this new setting. 
As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities. Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
%Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
%Please also note that, for our simple example, we do not have anymore to specify the all steps
%of a interactive system, as the communication mechanism from $rs2$ to $rs1$ will generate an infinite
%behaviour.
%\noindent
%First, we translate the rules in $rs1$:
First, we translate the reaction in $rs1$:
\[
\begin{array}{lcl}
%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\widetilde{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{r1_2}.P_{r1_1} +
%\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
\llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\widetilde{x}_o}\chainend{a_2}.P_{a_1} +
\startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
%\end{array}
%\]
%\\
%
%\begin{array}{lcl}
%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\widetilde{s1}}{\noact}\chainedlink{\noact}{\widetilde{s1}}\chainend{r2_1}.P_{r2_1} +
%\startchain{r1_2}\chainedlink{s2}{\noact}\chainedlink{\noact}{s2}\chainend{r2_1}.P_{r2_1}\\
\end{array}
\]

\noindent
Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
The translation for the entities follows.
\[
\begin{array}{lclcl}
\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
%\begin{array}{lclclll}
\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\\
%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{\widetilde{s2}}.P_{s2} +
%\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
%\\
%\\
&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\end{array}
%\]
%
%\[

@@

\noindent
The translation for $rs2$ follows.

\[
\begin{array}{lcl}
\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
\startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
\chainedlink{dss_i}{\noact}\chainedlink{\noact}{dss_o}\chainend{\silent}.P_{a_2}\\
%\end{array}
%\]
%
%\[
%\begin{array}{lcl}
%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\widetilde{s3}}{\noact}\chainedlink{\noact}{\widetilde{s3}}\chainend{\silent}.P_{r2_2}
%+
%\startchain{r2_2}\chainedlink{\overline{ss1}}{\noact}\chainedlink{\noact}{\overline{ss1}}\chainend{\silent}.P_{r2_2}\\
\end{array}
\]
\noindent
In the translation of the entities in  $rs2$, we introduce the mechanism that allows the entity
$s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$:
\[
\begin{array}{lclcl}
\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\

>>>>

% !TEX root = ./main_TCS.tex
%\section{Discussion}

@@

Our encoding increases the expressivity of  RS  concerning:
%the behaviour of the context,
the possibility of alternative behaviour of
mutated entities, and  the communication between two different 
RSs.
%Reaction Systems.
It is important to note that, when the context is deterministic, our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one state is reachable, 
as the dynamics is totally deterministic.

@@

\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline
$rs1$  & $rs2$ \\
\hline
%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
$a_1=(s,\ , x)$ & $a_2=(y,\  ,s)$\\
%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
\hline 
\end{tabular}
\caption{The two Reaction Systems $rs1$ and $rs2$.}
\label{tab:rules}
\end{table}
 \noindent
Now, we set our example such that the two contexts, for  $rs1$ and $rs2$, do not provide any entities.
We also assume that entity $s$ in $rs1$ is provided by $rs2$, as $rs2$ produces a quantity of $s$ that is enough for $rs1$ and $rs2$.
For technical reasons, we can not use the same name for  $s$ in both the two RSs, then we use the name $ss$ in $rs2$.
%
We need to modify our {\color{red} embedding}
%translation 
technique to suite this new setting. 
As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities. Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
%Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
%Please also note that, for our simple example, we do not have anymore to specify the all steps
%of a interactive system, as the communication mechanism from $rs2$ to $rs1$ will generate an infinite
%behaviour.
%\noindent
%First, we translate the rules in $rs1$:
First, we translate the reaction in $rs1$:
\[
\begin{array}{lcl}
%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\widetilde{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{r1_2}.P_{r1_1} +
%\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
\llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\widetilde{x}_o}\chainend{a_2}.P_{a_1} +
\startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
%\end{array}
%\]
%\\
%
%\begin{array}{lcl}
%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\widetilde{s1}}{\noact}\chainedlink{\noact}{\widetilde{s1}}\chainend{r2_1}.P_{r2_1} +
%\startchain{r1_2}\chainedlink{s2}{\noact}\chainedlink{\noact}{s2}\chainend{r2_1}.P_{r2_1}\\
\end{array}
\]

\noindent
Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
The {\color{red} encoding} 
%translation 
for the entities follows.
\[
\begin{array}{lclcl}
\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
%\begin{array}{lclclll}
\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\\
%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\widetilde{s2}}\chainend{\widetilde{s2}}.P_{s2} +
%\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
%\\
%\\
&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\end{array}
%\]
%
%\[

@@

\noindent
The {\color{red} encoding}
%translation 
for $rs2$ follows.

\[
\begin{array}{lcl}
\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
\startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
\chainedlink{dss_i}{\noact}\chainedlink{\noact}{dss_o}\chainend{\silent}.P_{a_2}\\
%\end{array}
%\]
%
%\[
%\begin{array}{lcl}
%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\widetilde{s3}}{\noact}\chainedlink{\noact}{\widetilde{s3}}\chainend{\silent}.P_{r2_2}
%+
%\startchain{r2_2}\chainedlink{\overline{ss1}}{\noact}\chainedlink{\noact}{\overline{ss1}}\chainend{\silent}.P_{r2_2}\\
\end{array}
\]
\noindent
In the {\color{red} encoding}
%translation 
of the entities in  $rs2$, we introduce the mechanism that allows the entity
$s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$:
\[
\begin{array}{lclcl}
\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\

### 2020-06-25 TCS_version/08-conclusion.tex
embedding encoding translation

In this paper we have introduced c\CNA, a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links, as it was 
briefly described in the future work section in~\cite{BBB17}.
This variant allowed us to define 
%shown 
a faithful embedding of 
RSs,
%Reaction Systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This translation shows several benefits.
For instance, in our paper the context of a RS can work non deterministically and it 
is recursively defined.
%  the context behaviour is expressed recursively and no be expressed recursively. 
Also, entity mutations can be expressed easily and different reaction systems can
communicate between them.

We have then modified the classical notion (in concurrency) of bisimulation
making it more suitable for proving properties of 
RSs
%Reaction Systems 
in our
framework.
We have defined a new assertion language, which allows us to specify
the properties to be verified, and have shown that it extends Hennessy-Milner logic
to our framework.
We believe that our work can make 
possible to investigate how to integrate our methodology
with other formal techniques to prove 
properties of the modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.

@@

As future work, we plan to implement a prototype of our framework,
with an automatic translation from RSs to  {\tt link}-calculus. 
We can also exploit the implementation of  the symbolic semantics of 
the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.

>>>>

In this paper we have introduced c\CNA, a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links, as it was 
briefly described in the future work section in~\cite{BBB17}.
This variant allowed us to define 
%shown 
a faithful embedding of 
RSs,
%Reaction Systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This {\color{red} embedding}
%translation 
shows several benefits.
For instance, in our paper the context of a RS can work non deterministically and it 
is recursively defined.
%  the context behaviour is expressed recursively and no be expressed recursively. 
Also, entity mutations can be expressed easily and different reaction systems can
communicate between them.

We have then modified the classical notion (in concurrency) of bisimulation
making it more suitable for proving properties of 
RSs
%Reaction Systems 
in our
%framework.
{\color{red} embedding.}
We have defined a new assertion language, which allows us to specify
the properties to be verified, and have shown that it extends Hennessy-Milner logic
to our 
%framework.
{\color{red} embedding.}
We believe that our work can make 
possible to investigate how to integrate our methodology
with other formal techniques to prove 
properties of the modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.

@@

As future work, we plan to implement a prototype of our 
%framework,
{\color{red} embedding,}
with an automatic translation from RSs to  {\tt link}-calculus. 
We can also exploit the implementation of  the symbolic semantics of 
the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.
